---
title: "Dimensionality reduction: fs_pca() and fs_svd()"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(featR)
```

`fs_pca()` and `fs_svd()` do **not** select features. They replace the
original columns with a few new ones, each a linear combination of *all* the
inputs. You still have to measure every variable, and the components are
chosen without reference to any outcome. The highest-variance directions need
not be the ones that predict anything.

Both functions return a plain list rather than an `fs_result`.

## `fs_pca()`

`fs_pca()` runs a PCA on the numeric columns of a data frame. Character and
factor columns are kept aside as labels. Incomplete rows and zero-variance
columns are dropped before the decomposition, and `meta` records what was used.

```{r}
pca <- fs_pca(iris, num_pc = 2, label_col = "Species")
names(pca)
pca$var_explained
pca$pc_loadings
head(pca$pca_df)
pca$meta$n_rows_used
```

### Reading the output

* `var_explained` is the share of the **total** variance each retained
  component carries. With two components out of four it sums to less than 1.
* `pc_loadings` holds the eigenvectors (`prcomp()`'s `rotation`), one row per
  input column. Signs are arbitrary. A component and its negation are the same
  component.
* `pc_scores` and `pca_df` hold the coordinates of each observation.

By default columns are centered **and scaled** (`scale_data = TRUE`), so the
PCA is on the correlation matrix. Leave that on unless the columns share
meaningful units. Otherwise the column with the largest variance dominates.

### Plotting

A ggplot of PC1 against PC2 is built whenever a `label_col` is supplied, two or
more components are kept, and ggplot2 is installed. `plot = TRUE` also prints
it:

```{r, eval = requireNamespace("ggplot2", quietly = TRUE), fig.width = 6, fig.height = 4}
pca$plot
```

### Large data

Below 10 million cells, `fs_pca()` uses `stats::prcomp()`. At or above that
size it uses `bigstatsr::big_SVD()` on a temporary file-backed matrix, if
bigstatsr is installed, and computes only the requested components. In that
case the total-variance denominator is computed from column statistics, so
`var_explained` still means share of the whole.

## `fs_svd()`

`fs_svd()` is the lower-level tool. It takes a numeric matrix `x` and returns
singular values and vectors:

```{r}
m <- as.matrix(mtcars)
s <- fs_svd(m, n_singular_values = 3, scale_input = TRUE)
round(s$singular_values, 3)
dim(s$left_singular_vectors)
dim(s$right_singular_vectors)
```

`scale_input = TRUE` (the default) centers and scales the columns first, so
the triplets factorize `scale(x)`, not `x`. The squared singular values
divided by `n - 1` are the PCA eigenvalues of the scaled data:

```{r}
all_sv <- fs_svd(m)$singular_values
round(all_sv^2 / (nrow(m) - 1), 4)[1:3]
round(prcomp(m, scale. = TRUE)$sdev^2, 4)[1:3]
```

### Exact or approximate

`svd_method = "auto"` uses `base::svd()` unless you ask for fewer singular
values than `min(dim(x))` **and** the matrix is larger than `svd_threshold`
(100) in its smaller dimension. In that case it uses `RSpectra::svds()`, which
computes only the leading triplets. `svd_method = "approx"` forces the
approximate solver when it is applicable.

```{r, eval = requireNamespace("RSpectra", quietly = TRUE)}
set.seed(1)
big <- matrix(rnorm(400 * 150), 400, 150)
approx <- fs_svd(big, n_singular_values = 5, svd_method = "approx")
exact  <- fs_svd(big, n_singular_values = 5, svd_method = "exact")
all.equal(approx$singular_values, exact$singular_values, tolerance = 1e-6)
```

Singular vectors are unique only up to sign, so compare them with
`abs()` or through reconstructions, not element by element.
