---
title: "Filter methods"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(featR)
```

Filters score each feature without fitting a predictive model. They are the
cheapest methods in featR and the right first step on wide data. They are also
**univariate**: each feature is judged on its own, so a filter cannot see
redundancy or interactions.

The examples below use one simulated data set, so the answers are known in
advance:

```{r}
set.seed(1)
n <- 200
sim <- data.frame(
  x_strong = rnorm(n),
  x_weak   = rnorm(n),
  x_noise  = rnorm(n),
  x_const  = 5,
  x_gappy  = ifelse(runif(n) < 0.6, NA, rnorm(n))
)
sim$x_copy <- sim$x_strong + rnorm(n, sd = 0.05)   # near-duplicate
sim$y      <- 2 * sim$x_strong + 0.5 * sim$x_weak + rnorm(n)
sim$class  <- factor(ifelse(sim$y > 0, "pos", "neg"))
```

## `fs_unsupervised()`: clean-up without an outcome

`fs_unsupervised()` scores each column by a target-free statistic: `"variance"`,
`"mad"`, `"iqr"`, `"range"`, `"missing_prop"`, or `"n_unique"`. Because it never
looks at the outcome, it is safe to run before a train/test split.

Drop constant columns:

```{r}
num <- sim[, c("x_strong", "x_weak", "x_noise", "x_const", "x_copy")]
fs_unsupervised(num, method = "variance", threshold = 0)
```

Drop columns that are more than half missing. Note `action = "remove"`. The
default `action = "keep"` would keep the *most* missing columns, and featR
warns if you leave all three of `threshold`, `direction`, and `action` at their
defaults for `"missing_prop"`.

```{r}
fs_unsupervised(sim[, c("x_strong", "x_gappy")], method = "missing_prop",
                threshold = 0.5, action = "remove")$selected
```

Thresholds are on the scale of the statistic: a variance is in squared units,
`"missing_prop"` is in [0, 1], and `"n_unique"` is a count. A threshold chosen
for one method means nothing for another. `"mad"` uses R's default
normal-consistency constant (1.4826), so it estimates a standard deviation.

## `fs_supervised()`: one feature at a time against the outcome

For a numeric target the score is the absolute Pearson correlation. For a
factor target it is the one-way ANOVA *F* statistic. `method = "auto"` picks
the right one.

```{r}
reg <- sim[, c("x_strong", "x_weak", "x_noise", "x_copy", "y")]
res <- fs_supervised(reg, target = "y", threshold = 0.2)
summary(res)
```

`x_copy` is kept alongside `x_strong` even though it adds nothing. This is the
univariate blind spot. The *F* scale for a factor target is unbounded, so pick
a threshold on that scale:

```{r}
cls <- sim[, c("x_strong", "x_weak", "x_noise", "class")]
fs_supervised(cls, target = "class", threshold = 10)$scores
```

Pearson correlation measures **linear** association only. A feature with a
strong U-shaped effect can score near zero. Use `fs_infogain()` or a
model-based method if you suspect non-linear effects.

## `fs_correlation()`: remove redundancy

`fs_correlation()` ignores the outcome and looks at how the features relate to
each other. With `prune = TRUE` (the default) it drops features one at a time
until no two survivors correlate above `threshold`:

```{r}
red <- fs_correlation(num[, c("x_strong", "x_weak", "x_noise", "x_copy")],
                      threshold = 0.9)
red$selected
red$details$dropped
red$details$pairs
```

Pruning uses the `caret::findCorrelation()` heuristic. Within the strongest
remaining pair, it drops the member with the larger mean absolute correlation
to everything else. That choice does not consult the outcome, so it may drop
the member that predicts better. `fs_boruta()` prunes by importance instead.

Supported correlation types are `"pearson"`, `"spearman"`, `"kendall"`,
`"polychoric"` (ordered factors, needs polycor), and `"pointbiserial"`
(continuous against dichotomous).

## `fs_chi()`: categorical features against a categorical outcome

`fs_chi()` runs a chi-squared test of independence for every factor or
character feature. It applies a multiple-testing correction (Bonferroni by
default) and selects features whose adjusted p-value is below `sig_level`.
When any expected cell count is below 5 it switches to a Monte-Carlo p-value,
so pass `seed` for reproducibility.

```{r}
cat_data <- data.frame(
  colour = factor(ifelse(sim$x_strong > 0, "red", "blue")),
  size   = factor(sample(c("S", "M", "L"), n, replace = TRUE)),
  class  = sim$class
)
chi <- fs_chi(cat_data, target = "class", seed = 1)
chi$details$results
```

Scores are adjusted p-values, so **smaller is stronger**. `summary()` sorts
them ascending. A p-value measures evidence against independence, not effect
size: with enough rows, a negligible association is still significant. Numeric
columns are ignored, so discretize them first if you want them tested.

## `fs_infogain()`: bits of uncertainty removed

Information gain is `H(Y) - H(Y | X)` in bits. Numeric predictors are
discretized into equal-width bins. A numeric target is discretized once,
up front, so that every predictor is scored against the same classes.

```{r}
ig <- fs_infogain(sim[, c("x_strong", "x_weak", "x_noise", "class")],
                  target = "class")
ig$scores
```

Raw information gain favors features with many levels. An ID column can
achieve the maximum possible gain while predicting nothing. Use
`normalize = "gain_ratio"`, which divides by the feature's own entropy, when
features differ in cardinality:

```{r}
with_id <- cbind(id = factor(seq_len(n)), sim[, c("x_strong", "class")])
fs_infogain(with_id, "class")$scores
fs_infogain(with_id, "class", normalize = "gain_ratio")$scores
```

The gain ratio cuts the ID column's score by roughly a factor of eight, but on
this data the ID column still ranks first. The gain ratio reduces the
cardinality bias. It does not remove it. Drop identifier and other
near-unique columns before you score anything.

By default `selected` holds every feature with a score above zero. With finite
data almost every feature has *some* positive sample gain, so use `top_n` for a
meaningful cut.
