---
title: "The fs_result object"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(featR)
```

Fourteen of featR's sixteen functions return the same object, an `fs_result`.
This article covers what is in it, how to read it, and how to use it to compare
methods. The two exceptions, `fs_pca()` and `fs_svd()`, do dimensionality
reduction and return their own lists. The
[dimensionality reduction](dimensionality-reduction.html) article covers them.

## Anatomy

An `fs_result` is a plain list with seven elements, always in this order:

| Element | Type | Meaning |
|---|---|---|
| `selected` | character | The features the method kept. It may be `character(0)`, never `NULL`. |
| `scores` | named numeric, data.frame, or `NULL` | Per-feature scores. They are comparable within one method, not across methods. |
| `method` | string | What actually ran, for example `"supervised_correlation"` or `"svm_linear"`. |
| `task` | string | `"classification"`, `"regression"`, or `NA`. |
| `model` | object or `NULL` | The fitted model, when the method fits one. |
| `details` | named list | Method-specific extras, documented on each help page. |
| `call` | call | The matched call. |

```{r}
d <- data.frame(
  signal = c(1.1, 2.3, 2.9, 4.2, 5.1, 5.8, 7.2, 8.1),
  flip   = c(8, 7, 6, 5, 4, 3, 2, 1),
  noise  = c(3, 1, 4, 1, 5, 9, 2, 6),
  y      = 1:8
)
res <- fs_supervised(d, target = "y", threshold = 0.5)

str(unclass(res), max.level = 1, give.attr = FALSE)
```

## Reading a result

`print()` gives a compact overview:

```{r}
res
```

`summary()` adds the call and a ranked score table. An asterisk marks the
features that were kept:

```{r}
summary(res)
```

`selected()` is the stable accessor for the chosen names. Prefer it to
`res$selected` in code you intend to keep:

```{r}
selected(res)
```

## What `scores` means, method by method

Scores are only comparable **within** one method. A correlation of 0.8 and a
Boruta importance of 0.8 have nothing in common. The direction also differs:
for `fs_chi()`, smaller is better. `summary()` knows this and sorts
accordingly.

| Function | `scores` | Better is |
|---|---|---|
| `fs_supervised()` | absolute Pearson *r* (numeric target) or ANOVA *F* (factor target) | larger |
| `fs_unsupervised()` | variance, MAD, IQR, range, missing proportion, or distinct count | depends on the criterion |
| `fs_chi()` | multiplicity-adjusted p-value | **smaller** |
| `fs_infogain()` | information gain in bits, or gain ratio | larger |
| `fs_correlation()` | each variable's largest absolute correlation with any other variable | larger means **more redundant** |
| `fs_lasso()` | data.frame of standardized coefficients (`AbsCoefficient` ranks) | larger magnitude |
| `fs_elastic()` | absolute coefficients on the scale the model saw (`NULL` for multinomial) | larger |
| `fs_randomforest()` | permutation importance (mean decrease in accuracy, or %IncMSE) | larger |
| `fs_mars()` | earth variable importance via `caret::varImp()` | larger |
| `fs_recursivefeature()` | resample-averaged importance from the RFE model | larger |
| `fs_svm()` | SVM-RFE criterion `w^2`, or random-forest importance (`NULL` without selection) | larger |
| `fs_boruta()` | median Boruta importance | larger |
| `fs_stepwise()` | absolute *t* statistics of the retained terms (not valid for inference) | larger |
| `fs_bayes()` | `NULL`: selection is between whole models, so there is no per-feature score | not applicable |

Some methods score **design-matrix columns** rather than original columns.
`fs_lasso()`, `fs_elastic()`, and `fs_svm()` all do this. A factor predictor
`colour` therefore shows up as `colourgreen`, `colourred`, and so on. Map the
names back yourself if you need the original variables.

## Comparing methods

Because every result has the same shape, a comparison is a loop:

```{r}
d2 <- data.frame(
  x1 = c(2.1, 3.9, 6.2, 7.8, 10.1, 12.2, 13.8, 16.1, 18.2, 19.9),
  x2 = c(1, 3, 2, 5, 4, 6, 8, 7, 10, 9),
  x3 = c(5, 5, 4, 5, 6, 5, 4, 5, 6, 5),
  y  = c(1.0, 2.1, 2.9, 4.2, 5.0, 6.1, 6.8, 8.1, 9.0, 9.9)
)

runs <- list(
  supervised = fs_supervised(d2, "y", threshold = 0.9),
  infogain   = fs_infogain(d2, "y", top_n = 2)
)
lapply(runs, selected)
```

Agreement between methods that work on different principles is weak evidence
that a feature matters. Disagreement tells you to look closer. Neither is a
substitute for validating the final set on data that played no part in
choosing it. The [validation and leakage](validation-and-leakage.html)
article covers that.

## Empty selections

When nothing passes, `selected` is `character(0)` and a warning says so. Code
that consumes results should handle a length-zero selection:

```{r}
none <- suppressWarnings(fs_unsupervised(d[, 1:3], threshold = 1e6))
selected(none)
length(selected(none)) == 0
```
