---
title: "Regularization and embedded importance"
---

```{r setup, include = FALSE}
# message = FALSE hides the "Loading required package" notes that caret
# prints when it attaches its model packages.
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE)
library(featR)
has <- function(...) all(vapply(c(...), requireNamespace, logical(1), quietly = TRUE))
```

These methods fit **one** model and read the selection off its structure:
non-zero coefficients (`fs_lasso()`, `fs_elastic()`), permutation importance
(`fs_randomforest()`), or the terms a spline model retains (`fs_mars()`).
Unlike filters, they judge each predictor in the presence of the others.

The examples use a simulated regression problem with a known answer.
`x1`–`x3` matter, `x1_copy` duplicates `x1`, `x4`–`x10` are noise, and `x_nl`
has a purely non-linear effect.

```{r}
set.seed(11)
n <- 300
d <- as.data.frame(matrix(rnorm(n * 10), n, 10,
                          dimnames = list(NULL, paste0("x", 1:10))))
d$x1_copy <- d$x1 + rnorm(n, sd = 0.1)
d$x_nl    <- runif(n, -2, 2)
d$y <- 2 * d$x1 - 1.5 * d$x2 + 0.75 * d$x3 + 1.5 * (d$x_nl^2 - 4 / 3) + rnorm(n)
```

## `fs_lasso()`

`fs_lasso()` fits an L1-penalized linear model with `glmnet::cv.glmnet()` and
keeps the predictors with non-zero coefficients at the cross-validated
penalty. It handles numeric outcomes only (gaussian family).

```{r, eval = has("glmnet", "Matrix")}
las <- fs_lasso(d, "y", nfolds = 5, seed = 1)
las
head(las$scores, 6)
```

Things to know:

* **Which penalty.** `lambda = "min"` (the default) uses the penalty that
  minimizes cross-validated error. That choice is tuned for prediction and
  usually keeps some noise variables. `lambda = "1se"` uses the largest
  penalty within one standard error of the minimum, which gives a sparser and
  more stable set.

```{r, eval = has("glmnet", "Matrix")}
selected(fs_lasso(d, "y", nfolds = 5, seed = 1, lambda = "1se"))
```

* **Correlated predictors.** The lasso tends to keep one member of a
  correlated group and zero out the rest. Here `x1` and `x1_copy` compete, and
  an absent feature is **not** evidence that it is unrelated to the outcome.
* **Linear only.** `x_nl` has a strong U-shaped effect but almost no linear
  one. The lasso gives it only a small coefficient, from whatever linear trend
  this sample happens to contain, and badly understates how much it matters.
  Compare the random forest and MARS results below.
* **Scores.** `scores` holds coefficients multiplied by each column's standard
  deviation. That makes them comparable across predictors measured in
  different units. The raw coefficients are in `details$coefficients`.
* **Factors** are expanded into dummy columns, and the selection names those
  columns (`colourred`, not `colour`).
* **Missing values** are an error by default. Mean-imputing on the full data
  before cross-validation leaks information between folds.

## `fs_elastic()`

`fs_elastic()` tunes both the mixing parameter `alpha` and the penalty
`lambda` by resampling, through `caret::train(method = "glmnet")`. It also
supports classification. When the winning `alpha` is below 1, the ridge part
of the penalty spreads weight across correlated predictors, so a correlated
group tends to survive together.

```{r, eval = has("caret", "glmnet", "Matrix")}
el <- fs_elastic(d, "y", alpha_seq = c(0.25, 0.5, 1), seed = 1)
el$details$best_alpha
selected(el)
```

The default grid is `seq(0.1, 1, by = 0.1)`. It deliberately excludes
`alpha = 0`, pure ridge, which never sets a coefficient to zero: if ridge won
the tuning, every predictor would be "selected". `scores` here are absolute
coefficients on the scale of the columns as supplied, so standardize the
predictors first if you want to compare them. `use_pca = TRUE` fits the model
on principal components that are recomputed inside every resample, and then
the selection names components (`PC1`, ...) rather than your columns.

## `fs_randomforest()`

`fs_randomforest()` splits the data, grows a forest on the training rows, and
reports **permutation importance** plus metrics on the held-out rows. A forest
*ranks* rather than *selects*. Unless you pass a `control$feature_select`
hook, `selected` is every predictor, ordered by importance.

```{r, eval = has("randomForest", "caret")}
rf <- fs_randomforest(d, "y", task = "regression",
                      control = list(ntree = 300), seed = 1)
head(rf$scores, 6)
rf$details$metrics
```

The forest finds `x_nl`, which the linear methods cannot. To turn the ranking
into a selection, pass a hook. It runs on the training rows only, so the
held-out metrics stay honest:

```{r, eval = has("randomForest", "caret")}
top5 <- function(dt) {
  keep <- c("x1", "x2", "x3", "x_nl", "x1_copy")  # any rule you like
  dt[, c(keep, "y"), with = FALSE]
}
rf_sel <- fs_randomforest(d, "y", task = "regression",
                          control = list(ntree = 300, feature_select = top5),
                          seed = 1)
selected(rf_sel)
```

`task` must be given explicitly, because a numeric column can be a class label.
Permutation importance splits credit between correlated predictors (see `x1`
and `x1_copy`) and favors predictors with many distinct values, so treat close
scores as ties.

## `fs_mars()`

`fs_mars()` fits multivariate adaptive regression splines (earth) through
caret with repeated cross-validation. It reports the predictors that the
pruned model retains, and it evaluates on a held-out split. MARS finds
non-linear effects and interactions on its own, using piecewise-linear hinge
functions.

```{r, eval = has("caret", "earth")}
ma <- fs_mars(d, "y", degree = 1:2, nprune = c(5, 10, 15),
              number = 3, repeats = 1, seed = 1)
selected(ma)
ma$details$metrics
ma$details$removed_predictors
```

Before fitting, `fs_mars()` drops near-zero-variance predictors and one member
of each pair correlated above `corr_cut` (default 0.95). It fits those filters
on the training rows only, and `details$removed_predictors` records what went.
Here one of `x1` and `x1_copy` is removed before the model ever sees it. The
filter does not look at the outcome, so which one goes is arbitrary. For classification,
class upsampling (when requested by caret's `sampling`) happens inside each
resample. Test-set ROC and PR AUC are reported for binary outcomes when pROC
and PRROC are installed.

The reported regression `R2` is caret's squared correlation between predictions
and observations, not `1 - SSE/SST`. A biased model can still score well on it,
so read it together with RMSE.
