---
title: "Reproducibility, randomness, and parallelism"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(featR)
```

## featR does not touch your random stream

No featR function calls `set.seed()` unless you pass `seed`. When you do, the
seed applies **for that call only**. Your previous RNG state is restored when
the function returns, even when it errors. Two consequences follow:

* Two calls with the same `seed` give the same answer.
* Calling featR with a seed does not change the random numbers your own code
  draws afterwards.

```{r}
d <- data.frame(a = sin(1:50), b = cos(1:50), c = sin(1:50) + 0.1 * cos(3:52))

set.seed(2024)
expected <- runif(3)

set.seed(2024)
res <- fs_correlation(d, threshold = 0.9, sample_frac = 0.5, seed = 1)
observed <- runif(3)

# The seeded call left the caller's stream exactly where it was
identical(expected, observed)
```

Without a `seed`, randomized methods draw from your global stream like any
other R function, so `set.seed()` before the call also works.

### Which functions use randomness

* **Train/test splits:** `fs_randomforest()`, `fs_mars()`,
  `fs_recursivefeature()`, `fs_svm()`.
* **Cross-validation folds:** `fs_lasso()`, `fs_elastic()`, `fs_mars()`,
  `fs_recursivefeature()`, `fs_svm()`.
* **Model fitting:** `fs_randomforest()` and `fs_boruta()` grow random forests.
* **Monte-Carlo p-values:** `fs_chi()`, when expected cell counts are small.
* **Row sampling:** `fs_correlation(sample_frac < 1)`.
* **Subset sampling:** `fs_bayes(sample_combinations = ...)`.
* **Deterministic:** `fs_supervised()`, `fs_unsupervised()`,
  `fs_infogain()`, `fs_stepwise()`, `fs_pca()`, and `fs_svd()` with the exact
  solver.

`fs_bayes()` is special. Its `seed` controls only which subsets are sampled.
To make the MCMC reproducible, pass `brm_args = list(seed = ...)` as well.

## Parallelism is opt-in and bounded

Every function runs on one core unless you ask for more. The rules are:

* Worker requests are **capped** at `parallel::detectCores()`. Asking for 64
  on a 4-core laptop gives 4.
* Clusters featR creates are **stopped when the call returns**, including on
  error.
* Parallel backends you registered yourself are restored afterwards.
* With a `seed`, parallel workers get reproducible L'Ecuyer-CMRG streams. Two
  seeded parallel runs therefore agree with each other. They need not agree
  with a seeded *sequential* run, because the workers draw from different
  streams.

```r
fs_randomforest(big_data, "y", task = "regression", n_cores = 4, seed = 1)
fs_lasso(big_data, "y", parallel = TRUE, n_cores = 4, seed = 1)
```

One caveat: when `fs_randomforest()` grows its forest on more than one worker,
the pieces are merged with `randomForest::combine()`. That drops the
out-of-bag error estimates, so `details$oob` is `NULL`. Use `n_cores = 1` if
you need the OOB numbers.

## Optional dependencies

The only hard dependencies are data.table, parallel, stats, utils, and withr.
Every modeling engine is in `Suggests` and is loaded only on the code path that
needs it. A missing package gives an actionable error:

```
Error: Packages 'glmnet', 'Matrix' are required for lasso feature selection
but not installed. Install with: install.packages(c("glmnet", "Matrix"))
```

To install everything at once:

```r
install.packages(c(
  "Boruta", "brms", "caret", "doParallel", "e1071", "earth", "foreach",
  "furrr", "future", "ggplot2", "glmnet", "kernlab", "loo", "MASS", "Matrix",
  "MLmetrics", "pbapply", "polycor", "pROC", "PRROC", "randomForest",
  "RSpectra", "bigstatsr"
))
```

`fs_bayes()` also needs a working C++ toolchain for Stan. See the
[brms installation notes](https://github.com/paul-buerkner/brms#how-do-i-install-brms).
