---
title: "Wrapper methods"
---

```{r setup, include = FALSE}
# message = FALSE hides the "Loading required package" notes that caret
# prints when it attaches its model packages.
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE)
library(featR)
has <- function(...) all(vapply(c(...), requireNamespace, logical(1), quietly = TRUE))
```

Wrappers refit a model over many candidate subsets and keep the subset that
model likes best. They are the most expensive methods in featR, and their
answer is specific to the model family they wrap. The search itself can overfit
too, so always validate the chosen set on rows the search never saw.

## `fs_recursivefeature()`: recursive feature elimination

`fs_recursivefeature()` wraps `caret::rfe()`. It first splits the data. On the
training rows it repeatedly fits a model, ranks the predictors, and drops the
weakest. Resampling chooses the subset size. It then evaluates the result on
the held-out rows.

```{r, eval = has("caret", "randomForest", "e1071")}
rfe <- fs_recursivefeature(
  iris, "Species",
  sizes = 1:4,
  rfe_control = list(method = "cv", number = 5),
  seed = 1
)
selected(rfe)
rfe$details$optimal_size
rfe$details$test_metrics         # held out: the honest estimate
rfe$details$resampling_results   # inside the training rows
```

By default the elimination uses random forests (`caret::rfFuncs`). Pass
another caret function set to match your model, for example
`rfe_control = list(method = "cv", number = 5, functions = caret::lmFuncs)` for
linear regression. `handle_categorical = TRUE` one-hot encodes factors with an
encoder fitted on the training rows only. `return_final_model = TRUE` also
trains a caret model (`model_method`) on the selected features.

Only `details$test_metrics` is a held-out estimate. `resampling_results` is
computed inside the training rows over the candidate sizes, and the size with
the best score is chosen from it, so the winning size's score is optimistic.

## `fs_svm()`: SVM-RFE

`fs_svm()` trains a support vector machine through caret and kernlab. It can
run **SVM-RFE** (Guyon et al., 2002) first. SVM-RFE fits a linear SVM on
centered and scaled training data, ranks features by their squared weight
`w^2`, drops the weakest, and refits, until one feature is left. The order in
which features were eliminated gives the ranking.

```{r, eval = has("caret", "kernlab", "e1071")}
sv <- fs_svm(
  iris, "Species",
  task = "classification",
  kernel = "linear",
  tune_grid = data.frame(C = c(0.1, 1)),
  feature_select = TRUE,
  select_method = "svm_rfe",
  nfolds = 5,
  seed = 1
)
selected(sv)
sv$details$selection$ranking
sv$details$performance$overall[c("Accuracy", "Kappa")]
```

Without `n_features`, the number of features to keep is chosen by
cross-validation over a short ladder of sizes (1, 2, 4, 8, ..., p). You must
state `task` explicitly. SVM-RFE needs `kernel = "linear"`, because the
ranking uses the primal weight vector, which exists only for a linear kernel.
For other kernels, use `select_method = "rf_rfe"`, which screens with random
forests instead. Selection always runs on the training split, and
`details$performance` comes from the test split. Selection works on the
**dummy-encoded** predictors, so factor levels are selected individually.

## `fs_boruta()`: all-relevant selection

Boruta asks a different question: which features carry *any* information about
the outcome? It adds "shadow" copies of every feature with their values
shuffled, grows random forests, and confirms each feature that beats the best
shadow significantly more often than chance. Redundant but informative
features are **all** confirmed. That is useful for understanding a data set,
and less useful when you want a compact model.

```{r, eval = has("Boruta")}
set.seed(3)
bd <- data.frame(
  signal  = rnorm(150),
  noise1  = rnorm(150),
  noise2  = rnorm(150)
)
bd$echo <- bd$signal + rnorm(150, sd = 0.2)
bd$y <- factor(ifelse(bd$signal + rnorm(150, sd = 0.5) > 0, "a", "b"))

bo <- fs_boruta(bd, "y", maxRuns = 50, cutoff_cor = NULL, seed = 1)
bo$details$decisions
selected(bo)

# featR addition: within each correlated group, keep the most important member
selected(fs_boruta(bd, "y", maxRuns = 50, cutoff_cor = 0.7, seed = 1))
```

`resolve_tentative = TRUE` (the default) settles features still undecided
after `maxRuns` with `Boruta::TentativeRoughFix()`. That is a heuristic, not
further evidence, so raise `maxRuns` if many features end up tentative.

## `fs_stepwise()`: AIC stepwise regression

`fs_stepwise()` runs `MASS::stepAIC()` on a linear regression. It is cheap,
deterministic, and returns an ordinary `lm`.

```{r, eval = has("MASS")}
st <- fs_stepwise(mtcars, "mpg", direction = "both")
selected(st)
st$scores
```

Pass `k = log(nrow(data))` to use BIC, which selects fewer terms. Stepwise
search is greedy and unstable: small changes to the data can change the
answer. The *t* statistics in `scores` and the p-values in
`details$coefficients` are computed after selection on the same data, so they
are biased toward significance. **Do not use them for inference.**

## `fs_bayes()`: Bayesian model comparison

`fs_bayes()` fits a brms model for every candidate predictor subset and
compares them by approximate leave-one-out cross-validation, using
`loo::loo_compare()`. By default it applies a one-standard-error rule: the
chosen subset is the smallest one whose expected log predictive density is
within one standard error of the best.

```r
res <- fs_bayes(
  mtcars, "mpg",
  predictors = c("wt", "hp", "qsec"),
  brm_args = list(chains = 2, iter = 1000, seed = 1),
  seed = 1
)
selected(res)
res$details$loo_comparison
```

This is the most expensive method in featR. With `p` candidates it fits
`2^p - 1` models, and each one compiles and samples a Stan program. Bound the
search with `max_comb_size` or `sample_combinations`. It needs brms, loo, and
a working C++ toolchain, which is why this example is not run here.
`details$mae` and `details$rmse` are in-sample, and the posterior of the
returned model is post-selection, so neither is an honest estimate on its own.

## References

Guyon, I., Weston, J., Barnhill, S. and Vapnik, V. (2002). Gene selection for
cancer classification using support vector machines. *Machine Learning* 46,
389–422. <https://doi.org/10.1023/A:1012487302797>

Kursa, M. B. and Rudnicki, W. R. (2010). Feature selection with the Boruta
package. *Journal of Statistical Software* 36(11).
<https://doi.org/10.18637/jss.v036.i11>

Vehtari, A., Gelman, A. and Gabry, J. (2017). Practical Bayesian model
evaluation using leave-one-out cross-validation and WAIC. *Statistics and
Computing* 27, 1413–1432. <https://doi.org/10.1007/s11222-016-9696-4>
